Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems

### 2024-06-15 final/thin-figures.tex
Reorder figure 2
Two column figure 3
Thin Figure 5

  % FIGURE 2
  \begin{figure*}[t]

    \begin{minipage}[b]{0.50\linewidth}
    {
        \centering 
        \includegraphics[width=\textwidth]{ide1.pdf}
    }
    \end{minipage}%
    %
    \begin{minipage}[b]{0.50\linewidth}
    {
        \centering
        \includegraphics[width=\textwidth]{ide2.pdf}
    }
    \end{minipage}%

    \caption{
        \label{fig-ide}
        Screenshots of Visual Studio Code showing
        tooltips (green arrows) and
        highlighted type errors (red arrows).
        On the left,
        we show that \texttt{a},
        a \texttt{Constant} assigned with \texttt{assign(...,\ constant=True)},
        can be used for \texttt{Variable b}'s initial condition.
        Instead,
        it is flagged as a type error (red underlining)
        when using \texttt{c}, a \texttt{Parameter},
        for \texttt{Variable d}'s initial condition,
        The IDE automatically recognizes \texttt{e} as an \texttt{Equation},
        and provides autocompletion of the \texttt{Model}'s components.
        A tooltip is shown when composing models (green arrow, right),
        which show the expected variables and their default values.
        The IDE highlights wrong names (\texttt{z} is not a name in \texttt{Model})
        and mismatched types (\texttt{x} is \texttt{Variable} and \texttt{a} must be a number or a \texttt{Constant})
    }

  % FIGURE 3
  \begin{figure*}[t]
    \begin{minipage}[t]{\linewidth}
      \begin{CodeInput}
      \begin{Highlighting}[]
      \KeywordTok{import} \ClassTok{numpy} \KeywordTok{as} \ClassTok{np}
      \KeywordTok{from} \ClassTok{poincare} \KeywordTok{import} \ClassTok{Simulator}

      \VariableTok{sim} \OperatorTok{=} \ClassTok{Simulator}\KeywordTok{(\ClassTok{Oscillator})}
      \VariableTok{df} \OperatorTok{=} \VariableTok{sim}.\FunctionTok{solve}\KeywordTok{(}\VariableTok{save\_at}\OperatorTok{=}\ClassTok{np}.\FunctionTok{linspace}\KeywordTok{(}\ValueTok{0}, \ValueTok{5}, \ValueTok{101}\KeywordTok{))}
      \end{Highlighting}
      \end{CodeInput}
    \end{minipage}%
    \newline
    \begin{minipage}[t]{0.45\linewidth}
      \begin{CodeInput}
      \begin{Highlighting}[]
      \VariableTok{df}.\FunctionTok{head}\KeywordTok{()}
      \end{Highlighting}
      \end{CodeInput}
      %
      \begin{CodeOutput}
          \begin{tabular}[]{@{}lll@{}}
              \toprule
              & x & v \\
              time & & \\
              \midrule
              0.00 & 1.000000 & 0.000000 \\
              0.05 & 0.998750 & -0.049979 \\
              0.10 & 0.995005 & -0.099835 \\
              0.15 & 0.988772 & -0.149440 \\
              0.20 & 0.980069 & -0.198672 \\
              \botrule
          \end{tabular}
      \end{CodeOutput}
    \end{minipage}%
    \hfill
    \begin{minipage}[t]{0.45\linewidth}
      \begin{CodeInput}
      \begin{Highlighting}[]
      \VariableTok{df}.\FunctionTok{plot}\KeywordTok{()}\OperatorTok{;}
      \end{Highlighting}
      \end{CodeInput}
      %
      \begin{CodeOutput}
        \includegraphics[width=\textwidth]{cell-output}
      \end{CodeOutput}
    \end{minipage}%

    \caption{\label{fig-sim}Simulation of the \texttt{Oscillator} system
    from Figure~\ref{fig-second-order}. The output is a
    \texttt{pandas.DataFrame} with a column for each variable and the time
    as index. It is inspected and plotted with the \texttt{pandas} methods.}
  \end{figure*}

@@

  % FIGURE 5
  \begin{figure*}[t]
    \centering

@@

    \VariableTok{model }\OperatorTok{=}\ClassTok{ sbml}.\FunctionTok{load}\KeywordTok{(}\StringTok{"repressilator.sbml"}\KeywordTok{)}  \CommentTok{\# from existing SBML file}
    \VariableTok{model }\OperatorTok{=}\ClassTok{ biomodels}.\FunctionTok{load}\KeywordTok{(}\StringTok{"BIOMD12"}\KeywordTok{)}  \CommentTok{\# from BioModels}
    \VariableTok{model}
    \end{Highlighting}
    \end{CodeInput}

    \begin{CodeOutput}
        \begin{tabular}{@{}lll@{}}
            \multicolumn{3}{l}{Elowitz2000 - Repressilator} \\
            \toprule
            type & total & names \\
            \midrule
            variables  &  6 & PX, PY, PZ, X, Y, Z \\
            parameters & 17 & cell, beta, alpha0, alpha, eff, n, KM, tau\_mRNA, tau\_prot, ... \\
            equations  & 12 & Reaction1, Reaction2, Reaction3, Reaction4, Reaction5, ... \\
            \botrule
        \end{tabular}
    \end{CodeOutput}

    \caption{Creation of a model from a local SBML file or one uploaded to BioModels.}
    \label{fig-simbio-io}
  \end{figure*}

>>>>

% FIGURE 2
\begin{figure}[t]
  \includegraphics[width=\columnwidth]{ide1.pdf}
  \includegraphics[width=\columnwidth]{ide2.pdf}
  \caption{
    \label{fig-ide}
    Screenshots of Visual Studio Code showing
    tooltips (green arrows) and
    highlighted type errors (red arrows).
    On the left,
    we show that \texttt{a},
    a \texttt{Constant} assigned with \texttt{assign(...,\ constant=True)},
    can be used for \texttt{Variable b}'s initial condition.
    Instead,
    it is flagged as a type error (red underlining)
    when using \texttt{c}, a \texttt{Parameter},
    for \texttt{Variable d}'s initial condition,
    The IDE automatically recognizes \texttt{e} as an \texttt{Equation},
    and provides autocompletion of the \texttt{Model}'s components.
    A tooltip is shown when composing models (green arrow, right),
    which show the expected variables and their default values.
    The IDE highlights wrong names (\texttt{z} is not a name in \texttt{Model})
    and mismatched types (\texttt{x} is \texttt{Variable} and \texttt{a} must be a number or a \texttt{Constant})
  }  
\end{figure}

% FIGURE 3
\begin{figure}[t]
  \begin{minipage}[t]{\columnwidth}
    \begin{CodeInput}
    \begin{Highlighting}[]
    \KeywordTok{import} \ClassTok{numpy} \KeywordTok{as} \ClassTok{np}
    \KeywordTok{from} \ClassTok{poincare} \KeywordTok{import} \ClassTok{Simulator}

    \VariableTok{sim} \OperatorTok{=} \ClassTok{Simulator}\KeywordTok{(\ClassTok{Oscillator})}
    \VariableTok{df} \OperatorTok{=} \VariableTok{sim}.\FunctionTok{solve}\KeywordTok{(}\VariableTok{save\_at}\OperatorTok{=}\ClassTok{np}.\FunctionTok{linspace}\KeywordTok{(}\ValueTok{0}, \ValueTok{5}, \ValueTok{101}\KeywordTok{))}
    \VariableTok{df}.\FunctionTok{head}\KeywordTok{()}
    \end{Highlighting}
    \end{CodeInput}
    \begin{CodeOutput}
        \begin{tabular}[]{@{}lll@{}}
            \toprule
            & x & v \\
            time & & \\
            \midrule
            0.00 & 1.000000 & 0.000000 \\
            0.05 & 0.998750 & -0.049979 \\
            0.10 & 0.995005 & -0.099835 \\
            0.15 & 0.988772 & -0.149440 \\
            0.20 & 0.980069 & -0.198672 \\
            \botrule
        \end{tabular}
    \end{CodeOutput}
  \end{minipage}%
  \hfill
  \begin{minipage}[t]{\columnwidth}
    \begin{CodeInput}
    \begin{Highlighting}[]
    \VariableTok{df}.\FunctionTok{plot}\KeywordTok{()}
    \end{Highlighting}
    \end{CodeInput}
    %
    \begin{CodeOutput}
      \includegraphics[width=\textwidth]{cell-output}
    \end{CodeOutput}
  \end{minipage}%

  \caption{\label{fig-sim}Simulation of the \texttt{Oscillator} system
  from Figure~\ref{fig-second-order}. The output is a
  \texttt{pandas.DataFrame} with a column for each variable and the time
  as index. It is inspected and plotted with the \texttt{pandas} methods.}
\end{figure}

@@

% FIGURE 5
\begin{figure}[t]
  \centering

@@

  \CommentTok{\# from existing SBML file}
  \VariableTok{model }\OperatorTok{=}\ClassTok{ sbml}.\FunctionTok{load}\KeywordTok{(}\StringTok{"repressilator.sbml"}\KeywordTok{)}
  \CommentTok{\# from BioModels}
  \VariableTok{model }\OperatorTok{=}\ClassTok{ biomodels}.\FunctionTok{load}\KeywordTok{(}\StringTok{"BIOMD12"}\KeywordTok{)}
  \VariableTok{model}
  \end{Highlighting}
  \end{CodeInput}

  \begin{CodeOutput}
      \begin{tabular}{@{}lll@{}}
          \multicolumn{3}{l}{Elowitz2000 - Repressilator} \\
          \toprule
          type & total & names \\
          \midrule
          variables  &  6 & PX, PY, PZ, X, Y, Z \\
          parameters & 17 & cell, beta, alpha0, alpha, eff, n, ... \\
          equations  & 12 & Reaction1, Reaction2, Reaction3, ... \\
          \botrule
      \end{tabular}
  \end{CodeOutput}

  \caption{Creation of a model from a local SBML file or one uploaded to BioModels.}
  \label{fig-simbio-io}
\end{figure}
